---
title: "F1 2018 Australia, Qualifying"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)
library(knitr)
stub='aus'
options(knitr.kable.NA = '')
opts_chunk$set(fig.path = paste0("images/f1_2018_",stub,"-"))
```



```{r}
library(f1djR)
gp="Australia"
qlaps= read.csv(paste0("~/Dropbox/various/",stub,"18_qualilaptimes.csv"))
cutoff=c(15,10,2)
```

```{r}
##Session classifications page
#xx=pageGrabber('http://www.fia.com/events/fia-formula-one-world-championship/season-2017/session-classifications-19')
qr=qualiResults.df(2018,1)

qr=quali_progression_ergast_tx(qr)
```

```{r}
#library(grid)
#library(gridExtra)
#promoter= function(g) {grid.arrange(g, nrow=1,
             ##top="My title", 
#             bottom = textGrob("To learn how to create this chart: f1datajunkie.com",
#                               gp = gpar(fontface=3, fontsize=9),
#                               hjust=1, x=1))}
```
```{r}
#qr=quali_progression_ergast_tx(ergastifyQualiClass(fiaSessionClassQualifyingTidy(fiaTableGrabber(xx,4))))
qrm=quali_progression_ergast_melt(qr)
qrtmp=qr
for (i in c(1,2,3)) {
  qrtmp[[paste0('Q',i)]]=ifelse(is.na(qrtmp[[paste0('Q',i)]]),'',
                                paste0(qrtmp[[paste0('Q',i)]],' (',qrtmp[[paste0('q',i,'pos')]],')'))
}
```

Drivers progressed through the qualifying session as follows:
```{r quali_slopegraph}
promoter_ergastf1dj(core_qualifying_rank_slopegraph(qr,qrm,spacer=0.21,reverse=T))
```


```{r}
 qsr=subset(qr, subset=(q1time<q2time))
 qsr['driverName']=factor(qsr[['driverName']])
```

```{r results='asis'}
cat(paste0(knit_child(text="Going from Q1 to Q2:\n",quiet=TRUE),'\n'))

if (length(levels(qsr$driverName))==0) {
  cat(paste0(knit_child(text='* all drivers improved on their best laptime from the previous session',quiet=TRUE),'\n'))
}

for (name in levels(qsr$driverName)) {
  q1time=qsr[qsr['driverName']==name,][['q1time']]
  q2time =qsr[qsr['driverName']==name,][['q2time']]
  text='* `r name` failed to improve his time, recording `r q2time` compared to `r q1time` (`r q2time-q1time`s slower)'
  t=paste0(knit_child(text=text,quiet=TRUE),'\n')
  cat(text=t)
}
```


```{r results='asis'}
qsr=subset(qr, subset=(q2time<q3time))
qsr['driverName']=factor(qsr[['driverName']])
cat(paste0(knit_child(text="Going from Q2 to Q3:\n",quiet=TRUE),'\n'))

if (length(levels(qsr$driverName))==0) {
  cat(paste0(knit_child(text='* all drivers improved on their best laptime from the previous session',quiet=TRUE),'\n'))
}

for (name in levels(qsr$driverName)) {
  text="* `r name` failed to improve his time, recording `r qsr[qsr['driverName']==name,][['q3time']]` compared to `r qsr[qsr['driverName']==name,][['q2time']]` (`r qsr[qsr['driverName']==name,][['q3time']]-qsr[qsr['driverName']==name,][['q2time']]`s slower)"
  cat(paste0(knit_child(text=text,quiet=TRUE),'\n')) 
}
```

Order change going from Q2 to Q3, along with laptime improvement (-ve is a better time in Q3)
```{r}
 txt='select qr1.driverName, (qr1.q3time-qr1.q2time) imp1,  qr2.driverName,(qr2.q3time-qr2.q2time) imp2  from qr as qr1 join qr as qr2 where qr1.driverName!=qr2.driverName and qr1.q2pos>qr2.q2pos and qr1.q3pos<qr2.q3pos'
library(sqldf)
 sqldf(txt)
```

```{r quali_time_slopegraph}
promoter(core_qualifying_time_slopegraph(qr,qrm,spacer=0.21))
```

The official session results were recorded as follows:
```{r}
qrtmp=rename(qrtmp, c("driverName"="DRIVER", "qspos"="POS"))
kable(subset(qrtmp,select=c('POS','DRIVER','Q1',	'Q2',	'Q3')), format='markdown',row.names = FALSE)
rm(qrtmp)
```


## Session Utilisation

Activity throughout the three qualification sessions took place as follows:

```{r}
#restart hack - for ITA 2017

```

```{r quali_utilisation,fig.width=10}
#qlapsa=rawLap_augment_laptimes(qlaps)
qlapsb=rawLap_augment_quali(qlaps)
qlapsb=qsessionOverride(qlapsb,0,1200,2600)
qlapsb=quali_purplePatch(qlapsb)
promoter_ergastf1dj(augmented_session_utilisation_chart(qlapsb,2,session=paste("F1",gp,"2018, Qualifying")))
```

## Q1 

```{r quali1_gridplot}
intercept= c(mean(qr[!is.na(qr["q1pos"]) & (qr["q1pos"]==10 | qr["q1pos"]==11),"q1time"]))
promoter_ergastf1dj(gridPlotTime(qr,"q1time",pos="q1pos", session="Q1",intercepts = intercept))
```

### Cut-off Time Evolution

The cut-off time in Q1 evolved as follows:

```{r quali1_cutoff}
promoter_ergastf1dj(quali_session_times_plots(qlapsb[qlapsb['qsession']==1,],1,FALSE,FALSE,TRUE,cutoff)+theme_bw())
```

## Q2

```{r quali2_gridplot}
intercept= c(mean(qr[!is.na(qr["q2pos"]) & (qr["q2pos"]==10 | qr["q2pos"]==11),"q2time"]))
promoter_ergastf1dj(gridPlotTime(qr,"q2time",pos="q2pos", session="Q2",intercepts = intercept))
```

### Cut-off Time Evolution

The cut-off time in Q2 evolved as follows:

```{r quali2_cutoff}
promoter(quali_session_times_plots(qlapsb[qlapsb['qsession']==2,],2,FALSE,FALSE,TRUE,cutoff)+theme_bw())
```

## Q3


```{r quali3_gridplot}
intercept= c(mean(qr[!is.na(qr["q3pos"]) & (qr["q3pos"]==2 | qr["q3pos"]==3),"q3time"]))
promoter_ergastf1dj(gridPlotTime(qr,"q3time",pos="q3pos", session="Q3",intercepts = intercept))
```

### Front Row Cut-off Time Evolution

The cut-off time in Q3 evolved as follows:

```{r quali3_cutoff}
promoter_ergastf1dj(quali_session_times_plots(qlapsb[qlapsb['qsession']>2,],3,FALSE,FALSE,TRUE,cutoff)+theme_bw())
```


---
 To learn how to wrangle F1 results and timing data to create charts like these, see [f1datajunkie.com](https://f1datajunkie.com) or [buy the *Wrangling F1 Data With R* book](https://leanpub.com/wranglingf1datawithr).
