---
title: "F1 2018 Australia, Race"
output:
  md_document:
    variant: gfm
---

# Race Report

F1 race report...

```{r setup, include=FALSE,message=FALSE,warning=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
opts_chunk$set(fig.path = "images/f1_2018_aus-")

library(f1djR)
```

```{r echo=FALSE}
library(grid)
library(gridExtra)
```

```{r}

year=2018
raceNum=1
gp='AUS'
#lapTimes=lapsData.df(year,1,con=dbConnect(RSQLite::SQLite(), './ergastdb13.sqlite'))
#raceId=dbGetQuery(f1,'SELECT raceId FROM races WHERE year="2012" AND round="1"')
lapTimes=lapsData.df(year,raceNum)

lapTimes=battlemap_encoder(lapTimes)
```


```{r}
results=resultsData.df(year,raceNum)
if(!"code" %in% colnames(results)) {
  results['code']=apply(results['driverId'],2,function(x) driverCodeMap(x))
}
resultstmp=plyr::rename(results, c("code"="DRIVER", "carNum"="CARNO.","grid"="GRID", 
                             "pos"="RANK", "status"="STATUS","laps"="LAPS"))
resultstmp['FASTLAP']=with(resultstmp, paste0(fastlapnum,': ',fastlaptime,' (',fastlaprank,')'))
resultstmp$FASTLAP[resultstmp$FASTLAP=='NA: NA (NA)']=''
kable(subset(resultstmp,select=c('RANK','CARNO.','DRIVER','GRID', 'STATUS', 'LAPS', 'FASTLAP')),
      format='markdown', row.names = FALSE)
rm(resultstmp)
```

```{r, include=FALSE}
#battlemapFull_byDriver(lapTimes,'HAM','2017 RUS - HAM Battlemap',c(-20,20))
template=c(paste0("```{r battlemap_",year,"_",gp,"_{{code}}}"),
           paste0("promoter(battlemapFull_byDriver(lapTimes,'{{code}}','", year, ' ', gp, " - {{code}} Battlemap',c(-20,20)))"),
           '```')

#The list needs to be sorted in race finish order
plist=intersect(c(c(results['code']$code),unique(lapTimes['code'])$code))
#c('HAM','VET', 'VES','RIC','RAI','BOT')
src = lapply(plist, function(code) knit_expand(text=c(template)))
```

`r knit(text = unlist(src))`

```{r track_position_chart}
promoter(track_position_chart(lapTimes,list(c("BOT","+"),c("RAI","*")))+theme_bw())
```


```{r raceHistory_chart}
promoter(raceHistory_chart(lapTimes)+theme_bw())
```

```{r}
#battlemapFull_byDriver(lapTimes,'HAM','2017 RUS - HAM Battlemap',c(-20,20))
#track_position_concordance(lapTimes,code)
template=c(paste0("```{r trackconcordance_",year,"_",gp,"_{{code}}}"),
           "promoter(track_position_concordance(lapTimes,'{{code}}'))",
           '```')
src = lapply(plist, function(code) knit_expand(text=c(template)))

```

`r knit(text = unlist(src))`


## Pit Stops

```{r}
pitstops=pitsData.df(year,raceNum )
pitstops['code']=apply(pitstops['driverId'],2,function(x) driverCodeMap(x))
```

```{r pithistory}
promoter(pitHistory(pitstops))
```

```{r pitlollipop}
pitLollipop(pitstops)
```


```{r pitStackedCumul}
promoter(race.plot.pits.cumulativeTime(pitstops))
```

```{r pitDodged}
promoter(race.plot.pits.dodged(pitstops))
```




## Circuit History

---

This website is unofficial and is not associated in any way with the Formula 1 companies. F1, FORMULA ONE, FORMULA 1, FIA FORMULA ONE WORLD CHAMPIONSHIP, GRAND PRIX and related marks are trade marks of Formula One Licensing B.V.
