---
title: "F1 Free Practice Template"
output:
  md_document:
    variant: gfm
params:
  psession: 1
  stub: 'bah'
  outdir: 'practice'
  rootdir: './reports'
  year: 2018
  country: 'Bahrain'
  laptimeFilePath: "~/Dropbox/wranglingf1datawithr/f1djR/scraping/bah18_p1laptimes.csv"
  
---

```{r dummy , include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE, echo=FALSE)
#If we also set echo=FALSE, the code will be hidden in the knitted report
library(knitr)


psession = params$psession
rootdir = params$rootdir
outdir = params$outdir
stub = params$stub
year = params$year
country = params$country
laptimeFilePath = params$laptimeFilePath

options(knitr.kable.NA = '')

#https://rgriff23.github.io/2017/04/25/how-to-knit-for-mysite.html
opts_knit$set(base.dir = paste(rootdir, outdir, sep='/'))
opts_chunk$set(fig.path = paste0("images/f1_",year,"_",stub,"_P",psession,"-"))
```


```{r init}
library(f1djR)
plaps= read.csv(paste0(laptimeFilePath))
plapsb=rawLap_augment_laptimes(plaps)
```

## Session Utilisation

Activity in Free Practice `r psession` took place as follows:


```{r utilisation,fig.width=10}
#qlapsa=rawLap_augment_laptimes(qlaps)

#qlapsb=qsessionOverride(qlapsb,0,1200,2600)
#qlapsb=quali_purplePatch(qlapsb)
promoter(augmented_session_utilisation_chart(plapsb,2,ordertype='ontrack',session=paste("F1",country,year,", Free Practice",psession)))
```


```{r}
#head(plapsb)
```


```{r}
stints=data.frame()
for (name in levels(plapsb$name)){
  dft=plapsb[plapsb$name==name,]
  dft=streaks(dft$stint)
  dft['name']=name
  dft=dft[c('name','start','end','l')]
  stints=rbind(stints,dft)
}
stints=ddply(stints,.(name),transform, stintNumber=1:length(l))
stints['name']=factor(stints$name)
#head(stints)
```

### Stints Summary

```{r}
psummary1 = ddply(stints, .(name), summarise,  total=sum(l), longest = max(l), nstints=max(stintNumber))
psummary1=psummary1[rev(order(psummary1$total,psummary1$longest)),]
names(psummary1) = c('Driver','Total laps','Longest run (laps)',"Stints")

kable(psummary1,format='markdown', row.names = FALSE)
```
```{r}
#psummary=merge(psummary1,psummary2[,c('name','driverbest','difftolead','difftoprev')],by.x='Driver',by.y='name')
#psummary
```
```{r results='asis'}
for (name in psummary1$Driver){
  text="* `r name` completed `r sum(abs(stints[stints['name']==name,]['l']))` laps over `r nrow(stints[stints['name']==name,])` stints, with a longest run of `r max(abs(stints[stints['name']==name,]['l']))` laps"
  cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
}
```

### Long Runs

Laptimes on long runs of 8 or more laps
```{r longruntimes,fig.width=10}
longruns=merge(stints[abs(stints['l'])>=8,],plapsb,by.x=c('name','stintNumber'),by.y=c('name','stint'))
longruns=arrange(longruns,name,lap)
g=ggplot(longruns[!longruns['outlap'] & !longruns['pit'],])
g=g+geom_line(aes(x=lapInStint,y=rawtime,group=stintNumber,colour=factor(stintNumber)))
g+facet_wrap(~code)+guides(colour=FALSE)
```

### Performance Summary

```{r}
#Get row for which driverbest is th smallest, by driver
#https://stackoverflow.com/a/42099608/454773
psummary2=ddply(plapsb[plapsb$pit==FALSE,], .(name), function(x) x[which.min(x$driverbest),])
psummary2 =psummary2[,c('name','number','driverbest')]
psummary2 = psummary2[order(psummary2$driverbest),]
psummary2$diff = c(0,diff(psummary2$driverbest))
psummary2$gap = cumsum(psummary2$diff)
names(psummary2)=c('Driver','Number','Fastest Lap', 'Diff','Gap')
kable(psummary2[,c('Driver','Number','Fastest Lap','Gap', 'Diff')], format='markdown', row.names = FALSE)
```



---

This website is unofficial and is not associated in any way with the Formula 1 companies. F1, FORMULA ONE, FORMULA 1, FIA FORMULA ONE WORLD CHAMPIONSHIP, GRAND PRIX and related marks are trade marks of Formula One Licensing B.V.
