---
title: "Quali Session Result template"
output:
  md_document:
    variant: gfm
    
params:
  session: 'quali'
  stub: 'generic'
  outdir: 'race'
  rootdir: './reports'
  year: 2018
  qr: NA
  qlapsb: NA
  override: NA
  gp: 'generic'
---

```{r q_ce_setup, echo=FALSE}
library(knitr)
library(f1djR)

knitr::opts_chunk$set(warning = FALSE, message=FALSE, echo=FALSE)

rootdir = params$rootdir
outdir = params$outdir
stub = params$stub
year = params$year
gp = params$gp
qr = params$qr
qlapsb = params$qlapsb
override = params$override

options(knitr.kable.NA = '')
```


```{r qr_ce_opts, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
opts_knit$set(base.dir = paste(rootdir, outdir, sep='/'))
knitr::opts_chunk$set(fig.path = paste0("images/f1_" ,year, "_", stub, "-"))
```


```{r cutofflooper}
template=c("
## Q{{session}}",
"```{r quali{{session}}_gridplot}
c = ifelse({{session}}==3,2,10)
intercept= c(mean(qr[!is.na(qr['q{{session}}pos']) & (qr['q{{session}}pos']==c | qr['q{{session}}pos']==c+1),'q1time']))
promoter_ergastf1dj(gridPlotTime(qr,'q{{session}}time',pos='q{{session}}pos', session='Q{{session}}',intercepts = intercept))
```",
"
### `r ifelse({{session}}==3,'Front Row ','')`Cut-off Time Evolution

The cut-off time in Q{{session}} evolved as follows:",
"```{r quali{{session}}_cutoff}
promoter_ergastf1dj(quali_session_times_plots(qlapsb[qlapsb['qsession']=={{session}},],{{session}},FALSE,FALSE,TRUE,cutoff)+theme_bw())
```")
src = lapply(c(1,2,3), function(session) knit_expand(text=c(template)))

```
`r knit(text = unlist(src))`
