---
title: "F1 Story So Far - Championship Standings Template"
output:
  md_document:
    variant: gfm
    
params:
  stub: 'generic'
  outdir: 'reports'
  rootdir: './reports'
  cct: NA
  ergastdb: NA
  year: 2018
  round: 2
---
```{r ssf_cs_setup, include=FALSE}
library(knitr)

library(toOrdinal)

year=params$year
stub=params$stub
round=params$round
ergastdb=params$ergastdb

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

```

### Drivers' Championship - Standings


```{r cs_df_driver_champ_table}
q=paste0('SELECT position AS Position, d.code AS Driver, ds.points AS Points FROM driverStandings ds JOIN races r JOIN drivers d WHERE ds.raceId=r.raceId AND d.driverId=ds.driverId AND year=', year, ' AND round=', round-1, ' ORDER BY position ASC')
df_championship_at_round = dbGetQuery(ergastdb, q)
```

```{r for_tracery }
#for tracery, eg: coming in to the Nth round; if round<=3: after the first N round(s); if N-round <=3: with [only? another|just]+ N rounds to go; if half way etc. 
#?? a reg exp like parser fo generating, rather than matching, sentences?
```

Coming in to the `r toOrdinal(round)` round,  `r df_championship_at_round[df_championship_at_round['Position']==1,'Driver']` leads the Drivers' Championship on `r df_championship_at_round[df_championship_at_round['Position']==1,'Points']` points, `r df_championship_at_round[df_championship_at_round['Position']==1,'Points']-df_championship_at_round[df_championship_at_round['Position']==2,'Points']` points ahead of second placed `r df_championship_at_round[df_championship_at_round['Position']==2,'Driver']` on `r df_championship_at_round[df_championship_at_round['Position']==2,'Points']`  points. `r df_championship_at_round[df_championship_at_round['Position']==3,'Driver']` completes the top three on  `r df_championship_at_round[df_championship_at_round['Position']==3,'Points']` points.

```{r cs_df_driver_champ,results='asis'}
kable(df_championship_at_round, format='markdown',row.names = FALSE,encoding='UTF-8')
```


### Constructors' Championship - Standings
```{r cs_const_df_driver_champ_table}
q=paste0('SELECT position AS Position, c.name AS Constructor, cs.points AS Points FROM constructorStandings cs JOIN races r JOIN constructors c WHERE cs.raceId=r.raceId AND c.constructorId=cs.constructorId AND year=', year, ' AND round=', round-1, ' ORDER BY position ASC')
df_const_championship_at_round = dbGetQuery(ergastdb, q)
```

In the Constructors' Championship,  `r df_const_championship_at_round[df_const_championship_at_round['Position']==1,'Constructor']` lead the championship on `r df_const_championship_at_round[df_const_championship_at_round['Position']==1,'Points']` points, with `r df_const_championship_at_round[df_const_championship_at_round['Position']==2,'Constructor']` `r df_const_championship_at_round[df_const_championship_at_round['Position']==1,'Points']-df_const_championship_at_round[df_const_championship_at_round['Position']==2,'Points']` points behind in second.

```{r cs_const_df_driver_champ}
kable(df_const_championship_at_round, format='markdown',row.names = FALSE)
```


