---
title: "F1 Story So Far Template"
output:
  md_document:
    variant: gfm
params:
  round: 2
  stub: 'generic'
  outdir: 'storysofar'
  rootdir: './reports'
  year: 2018
---

```{r setup, include=FALSE}
library(knitr)
library(f1djR)
library(toOrdinal)
library(plyr)

round=params$round
year=params$year

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

## The Story So Far

Coming into the `r toOrdinal(round)` race of `r year`, the championship points race is as follows:

```{r echo = FALSE}
library(RMySQL)
ergastdb=dbConnect(MySQL(),user='root',password='f1',host='127.0.0.1',port=3399,dbname='ergastdb')
```

### Drivers' Championship - Standings
```{r echo = FALSE}
q=paste0('SELECT position AS Position, d.code AS Driver, ds.points AS Points FROM driverStandings ds JOIN races r JOIN drivers d WHERE ds.raceId=r.raceId AND d.driverId=ds.driverId AND year=', year, ' AND round=', round-1, ' ORDER BY position ASC')
df_championship_at_round = dbGetQuery(ergastdb, q)
kable(df_championship_at_round, format='markdown',row.names = FALSE)
```

### Constructors' Championship - Standings
```{r}
q=paste0('SELECT position AS Position, c.name AS Constructor, cs.points AS Points FROM constructorStandings cs JOIN races r JOIN constructors c WHERE cs.raceId=r.raceId AND c.constructorId=cs.constructorId AND year=', year, ' AND round=', round-1, ' ORDER BY position ASC')
df_const_championship_at_round = dbGetQuery(ergastdb, q)
kable(df_const_championship_at_round, format='markdown',row.names = FALSE)
```


```{r}
q=paste0('SELECT d.code, round, position, ds.points FROM driverStandings ds JOIN races r JOIN drivers d WHERE ds.raceId=r.raceId AND d.driverId=ds.driverId AND year=',year)
df_championship_by_round = dbGetQuery(ergastdb, q)
#df_championship_by_round
```



```{r}
library(ggplot2)
library(directlabels)
championship_race_charter=function(g) {
  #Generate a line chart
  g = g + geom_line()
  #Remove axis labels and colour legend
  g = g + ylab(NULL) + xlab(NULL) + guides(color=FALSE)
  #Add a title
  g = g + ggtitle("F1 Drivers' Championship Race, 2018")
  #Add right hand side padding to the chart so the labels don't overflow
  g + scale_x_continuous(limits=c(1,22), breaks=c(5,10,15))
}
```



```{r pointsRaceNonzero, message=FALSE,fig.cap="Cumulative points scored at the end of each round in the 2014 Drivers' Championship for drivers scoring at least 10 points by round 17"}
if (FALSE) {
  df.tenUp=df[df$round==(round-1) & df$points>=10,]$code

  #All the other elements of the chart definition are the same
  g=ggplot(df[df$code %in% df.tenUp,],aes(x=round,y=points,colour=code))
  g= championship_race_charter(g) 
  direct.label(g,list('last.bumpup',cex=0.7,dl.trans(x=x+0.2)))
}
```


```{r pointsRace10th,fig.cap="",message=FALSE}
if (FALSE){
  #This highlights a battle
  
  #Generate a guide that is the mean points value of 10th and 11th positions
  dfx=ddply(df[df['position']==10 | df['position']==11,],
            .(round), summarize, points=mean(points))
  #Set the driverId of this line to an empty string
  dfx$code=''
  
  #Identify the drivers fighting for 10th at the end of round 10
  #Include other drivers contending this position earlier in the season
  #Ideally, we would automatically identify the positions around this cluster
  df.contending = df[df$round==10 & df$pos>=9 & df$pos<=12,'driverId']
  #The following filter identifies drivers fighting for 10th after 17
  ##df.contending =df[df$round==17 & df$pos>=10 & df$pos<=12,'driverId']
  df.battle=df[df$driverId %in% df.contending,]
  
  #Base chart
  g=ggplot(df.battle,aes(x=round,y=points,colour=code))
  g=championship_race_charter(g)
  g=direct.label(g,list('last.bumpup',cex=0.7,dl.trans(x=x+0.2)))
  #Add in the guide line
  g = g + geom_line(data=dfx, aes(x=round, y=points),
                    col='black', linetype="dashed")
  g
}
```



```{r champRaceStandings,fig.cap="The championship race in terms of standings"}
#The lap chart style chart - plotting position rather than points
if (round>2){
  g = ggplot(df, aes(x=round, y=position,group=code))
  g = championship_race_charter(g)
  #If we know the size of the labels, we can easily fashion our own placement
  g = g + geom_text(data=df[df['round']==max(df['round']),],
              aes(x=round+0.9,label=code),size=3, vjust=-0.4)
  g = g+theme_bw()
  g
}
```



```{r champRaceStandingsTotPoints,fig.cap="Annotating the championship standings chart with championship points at each round, coloured by points earned per race"}
if (round>2){
  #Sort by ascending round
  df = arrange(df, round)
  #Derive how many points each driver scored in each race
  df = ddply(df, .(code), transform, racepoints=diff(c(0,points)))

  g = g + geom_text(data=df, aes(label=points, col=racepoints), vjust=-0.4, size=3)
  #Colour scheme for the points scored in each race
  g + scale_color_continuous(high='red')
}
```

```{r}
q=paste0('SELECT code FROM drivers d JOIN races r JOIN results rs WHERE d.driverId=rs.driverId AND r.raceId=rs.raceId AND year=',year)
driversThisYear= as.list(dbGetQuery(ergastdb, q))$code
driversThisYear_str = paste("'",driversThisYear,"'", collapse=',', sep='')
```

```{r}
q=paste0('SELECT c.name FROM constructors c JOIN races r JOIN results rs WHERE c.constructorId=rs.constructorId AND r.raceId=rs.raceId AND year=',year)
teamsThisYear= as.list(dbGetQuery(ergastdb, q))$name
teamsThisYear_str = paste("'",teamsThisYear,"'", collapse=',', sep='')
```

```{r}
#driverteamsthisyear
q_driverteamsthisyear=paste0('SELECT c.name name, d.code code FROM constructors c JOIN races r JOIN results rs JOIN drivers d WHERE c.constructorId=rs.constructorId AND r.raceId=rs.raceId and d.driverId=rs.driverId AND r.year=',year)
#dbGetQuery(ergastdb, q_driverteamsthisyear)
```

### Races to Date

```{r results='asis'}
#races to date modulo 25
q=paste0('SELECT d.code, COUNT(code)+1 AS races FROM drivers d JOIN results rs WHERE d.driverId=rs.driverId AND code IN (',driversThisYear_str,') GROUP BY code HAVING races>1 AND races % 25 = 0')
raceStartsThisTime  = dbGetQuery(ergastdb, q)

if (nrow(raceStartsThisTime)>0) {
  cat('This will be:\n')
  for (row in 1:nrow(raceStartsThisTime)) {
    text = " * `r raceStartsThisTime[row, 'code']`'s `r toOrdinal(raceStartsThisTime[row, 'races'])` grand prix"
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }

}
```


### Upcoming Race Trivia...


```{r results='asis'}

#These currently rely on db being current up to end of previous race
#need to add: AND (year<',year,' OR (year=',year,' AND round<',round,'))

upto=paste0(' AND (year<',year,' OR (year=',year,' AND round<',round,')) ')

## STARTS

#starts for a team
q=paste0('SELECT d.code, COUNT(code) AS startsforteam, c.name AS name FROM drivers d JOIN races r JOIN results rs JOIN constructors c WHERE c.constructorId=rs.constructorId AND d.driverId=rs.driverId AND r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') ',upto,' GROUP BY d.code, c.name HAVING (startsforteam+1) % 50 = 0')
startsTeammod50=dbGetQuery(ergastdb, q)

#looking for poles to date modulo 5 
q=paste0('SELECT d.code, COUNT(code) AS poles FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.grid=1',upto,' GROUP BY code HAVING poles>1 AND (poles+1) % 5 = 0')
lookingpolesmod5=dbGetQuery(ergastdb, q)

#pole vs bes qualifying - pole may be result of penalty to best qualifier
# TO DO

#looking for front row starts to date modulo 5 
q=paste0('SELECT d.code, COUNT(code) AS frontRows FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.grid IN (1,2) ',upto,' GROUP BY code HAVING frontRows>1 AND (frontRows+1) % 5 = 0')
frontrowmod5=dbGetQuery(ergastdb, q)

#looking for podiums to date modulo 5 
q=paste0('SELECT d.code, COUNT(code) AS podiums FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND  d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.position IN (1,2,3) ',upto,' GROUP BY code HAVING podiums>1 AND (podiums+1) % 5 = 0')
lookingpodiumsmod5=dbGetQuery(ergastdb, q)

#looking for wins to date modulo 5 
q=paste0('SELECT d.code, COUNT(code) AS wins FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.position=1 ',upto,' GROUP BY code HAVING wins>1 AND (wins+1) % 5 = 0')
lookingwinsmod5=dbGetQuery(ergastdb, q)

#Looking for wins from pole
q=paste0('SELECT d.code, COUNT(*) AS winsFromPole FROM drivers d JOIN results rs JOIN races r JOIN qualifying q WHERE r.raceId=rs.raceId AND d.driverId=rs.driverId AND q.driverId=rs.driverId AND q.raceId=rs.raceId AND q.position=1 AND rs.position=1 AND d.code IN (',driversThisYear_str,') ',upto,' GROUP BY d.code HAVING winsFromPole>1 AND (winsFromPole+1) % 10 = 0 ORDER BY winsFromPole DESC')
lookingwinspole10=dbGetQuery(ergastdb, q)

#wins for a teams
q=paste0('SELECT d.code, COUNT(*) AS winsfordriverteam, c.name AS name FROM drivers d JOIN races r JOIN results rs JOIN constructors c JOIN (',q_driverteamsthisyear,') dt WHERE c.constructorId=rs.constructorId AND d.driverId=rs.driverId AND r.raceId=rs.raceId AND rs.position=1 AND dt.code=d.code AND dt.name=c.name AND d.code IN (',driversThisYear_str,') ',upto,' GROUP BY d.code, c.name HAVING (winsfordriverteam+1) % 10 = 0')
lookingWinsDriverTeammod10=dbGetQuery(ergastdb, q)

if ((nrow(lookingpolesmod5) + nrow(lookingpolesmod5) + nrow(lookingpodiumsmod5) + nrow(lookingwinsmod5) +nrow(lookingwinspole10) + nrow(startsTeammod50) +nrow(lookingWinsDriverTeammod10))>0) {
  cat("Coming in to the race:\n")
}


if (nrow(startsTeammod50)>0) {
  for (row in 1:nrow(startsTeammod50)) {
    text = '- `r startsTeammod50[row, "code"]` is looking for their `r toOrdinal(startsTeammod50[row, "startsforteam"]+1)` start for `r startsTeammod50[row, "name"]`'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}

if (nrow(lookingpolesmod5)>0) {
  for (row in 1:nrow(lookingpolesmod5)) {
    text = '- `r lookingpolesmod5[row, "code"]` is looking for their `r toOrdinal(lookingpolesmod5[row, "poles"]+1)` ever pole position'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}


if (nrow(frontrowmod5)>0) {
  for (row in 1:nrow(frontrowmod5)) {
    text = '- `r frontrowmod5[row, "code"]` is looking for their `r toOrdinal(frontrowmod5[row, "frontRows"]+1)` ever front row start'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}


if (nrow(lookingpodiumsmod5)>0) {
  for (row in 1:nrow(lookingpodiumsmod5)) {
    text = '- `r lookingpodiumsmod5[row, "code"]` is looking for their `r toOrdinal(lookingpodiumsmod5[row, "podiums"]+1)` ever podium finish'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}


if (nrow(lookingwinsmod5)>0) {
  for (row in 1:nrow(lookingwinsmod5)) {
    text = '- `r lookingwinsmod5[row, "code"]` is looking for their `r toOrdinal(lookingwinsmod5[row, "wins"]+1)` ever race win'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}

lookingWinsDriverTeammod10
if (nrow(lookingWinsDriverTeammod10)>0) {
  for (row in 1:nrow(lookingWinsDriverTeammod10)) {
    text = '- `r lookingWinsDriverTeammod10[row, "code"]` is looking for their `r toOrdinal(lookingWinsDriverTeammod10[row, "winsfordriverteam"]+1)` ever win for `r lookingWinsDriverTeammod10[row,"name"]`'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}

if (nrow(lookingwinspole10)>0) {
  for (row in 1:nrow(lookingwinspole10)) {
    text = '- `r lookingwinspole10[row, "code"]` is looking for their `r toOrdinal(lookingwinspole10[row, "winsFromPole"]+1)` ever win from pole'
    cat(paste0(knit_child(text=text,quiet=TRUE),'\n'))
  }
}

```





```{r}
#TO DO
#consecutive finishes in points, on podium, in lead, pole, this season, this team, ever
```

```{r}
# TO DO
#Most wins by nationality
#wins by current drivers
q=paste0('SELECT d.driverRef, nationality, COUNT(*) AS wins FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.position=1 ',upto,' GROUP BY d.driverId, nationality')
#wins by nationality
q=paste0('SELECT d.driverRef, d.nationality, COUNT(*) AS wins FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.driverId=rs.driverId AND rs.position=1 ',upto,' GROUP BY d.driverId, nationality')
q=paste0('SELECT curr.driverRef AS Current, ever.driverRef AS AllTime,curr.nationality, ever.wins FROM (SELECT d.driverRef, nationality, COUNT(*) AS wins FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.code IN (',driversThisYear_str,') AND d.driverId=rs.driverId AND rs.position=1 ',upto,' GROUP BY d.driverId, nationality) curr, (SELECT d.driverRef, d.nationality, COUNT(*) AS wins FROM drivers d JOIN results rs JOIN races r WHERE r.raceId=rs.raceId AND d.driverId=rs.driverId AND rs.position=1 ',upto,' GROUP BY d.driverId, nationality) ever WHERE curr.nationality=ever.nationality and curr.wins=ever.wins and curr.driverRef!= ever.driverRef' )

#dbGetQuery(ergastdb, q)
```

```{r}
# TO DO
#points scoring streak
#pole streak
#podium streak
#front row streak
#winning streak

#celebrations
#eg if there is a 200 starts driver, what other drivers have had 200 starts (do a check on 0, <=5, <=10, <20; Nth driver to achieve it )
```

```{r}
## TO DO
#record - most wins from pole ever 
# - we don't have qualifying results far enough back in ergast db for this
q=paste0('SELECT d.driverRef, COUNT(*) AS winsFromPole FROM drivers d JOIN results rs JOIN races r JOIN qualifying q WHERE r.raceId=rs.raceId AND d.driverId=rs.driverId AND q.driverId=rs.driverId AND  q.raceId=rs.raceId AND q.position=1 AND rs.position=1 ',upto,' GROUP BY d.driverRef ORDER BY winsFromPole DESC')
#dbGetQuery(ergastdb, q)
```

### Circuit Performance

How well do drivers and constructors perform at the circuit?

The *Driver Performance Chart* and *Constructor Performance Chart* show how drivers and constructors fare based on points scored at the circuit. (Note that points schemes may change over the reporting period.)

```{r}
q=paste0('SELECT circuitRef, c.name AS circuit, r.name as race, location, country FROM races r JOIN circuits c WHERE r.circuitId=c.circuitId AND year=',year,' AND round=',round)
cct = as.list(dbGetQuery(ergastdb, q))
```


```{r}
library(plyr)
library(ggplot2)
library(directlabels)
library(ggrepel)
library(ggthemes)
library(reshape2)
```

```{r}
#Correlation betwen grid and finish position
q=paste0('SELECT grid, position, COUNT(*) AS Count FROM races r JOIN circuits c JOIN results rs WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND circuitref="',cct$circuitRef,'" GROUP BY grid, position DESC')
grid2pos =dbGetQuery(ergastdb, q)

g=ggplot(data=grid2pos, aes(x=grid, y=position))+geom_point(aes(size=Count))
g+geom_smooth(method = "lm", se = TRUE) +geom_abline(color='grey', linetype='dashed')+theme_bw()+ylab('Final Position')+xlab('Grid Position')
```

Historical likelihood of winning from pole: `r round(100*grid2pos[(grid2pos$grid==1) & (grid2pos$position==1),'Count']/sum(grid2pos[(grid2pos$position==1),'Count'], na.rm = TRUE), 2)`%



```{r}

champChartTheme=function(g){
  g = g + theme_minimal(base_family="Arial Narrow")
  #g = g + theme(panel.grid.major.y=element_blank())
  g = g + theme(panel.grid.minor=element_blank())
  g = g + theme(axis.line.y=element_line(color="#2b2b2b", size=0.15))
  g = g + theme(axis.text.y=element_text(margin=margin(r=0, l=0)))
  g = g + theme(plot.margin=unit(rep(30, 4), "pt"))
  g = g + theme(plot.title=element_text(face="bold"))
  g = g + theme(plot.subtitle=element_text(margin=margin(b=10)))
  g = g + theme(plot.caption=element_text(size=8, margin=margin(t=10)))
  g
}

champ_chart = function(res,leaders, typ, ylabel='Championship Points'){
  g=ggplot(data=res[res$code %in% leaders$code,], aes(x=year, y=totPts))+ geom_line(aes(col=code))
  g=g+geom_point(data=res[(res$position <=3) & (res$code %in% leaders$code),],aes( size=factor(position),shape=factor(position),col=code)) + scale_size_manual(values=c(2,1.2,1.0)) + scale_shape_manual(values=c(2,6,1))
  
  g=g+ geom_line(data=res[!(res$code %in% leaders$code),], aes(group=code),col='grey')
  
  #Label the lines
  g=g+geom_label_repel(data=leaders,aes(x=until, y=s,label = code, col=code), na.rm = TRUE, arrow = arrow(length = unit(0.01, 'npc')), nudge_y =3, nudge_x =2, segment.size = 0.2,size=2)
  
  g=g + guides(color=FALSE)+ylim(0,round_any(max(res$totPts)+10,10))#+theme_bw()
  g=g+xlim(min(res$year),max(res$year)+5)
  g = g+ylab(ylabel)+xlab(NULL)+scale_y_continuous(breaks=seq(0,round_any(max(res$totPts)+10,10),10))
  g=g + guides(size=FALSE, shape=FALSE)
  g=g+labs(x=NULL,
           title=paste0('F1 ', cct$race, ' - ',typ,', ', min(res$year), ' to ', max(res$year)),
           subtitle = paste0(cct$circuit,', ',cct$location,', ',cct$country),
           caption="Data from Ergast database, ergast.com/mrd")
  champChartTheme(g)
}
```



```{r driver_circuit_performance}

q=paste0('SELECT code, MAX(year) as until, SUM(points) s FROM races r JOIN circuits c JOIN results rs JOIN drivers d WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND  d.driverId=rs.driverId AND circuitref="' ,cct$circuitRef,'" GROUP BY code DESC ORDER BY s DESC LIMIT 10')
leaders = dbGetQuery(ergastdb, q)

q=paste0('SELECT year, code, position, points FROM races r JOIN circuits c JOIN results rs JOIN drivers d WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND d.driverId=rs.driverId AND circuitref="',cct$circuitRef,'" ORDER BY year')
 #The dbGetQuery() function is provided by the DBI library
res = dbGetQuery(ergastdb, q)

res = ddply(res, .(code), transform, totPts=cumsum(points)  )

champ_chart(res,leaders, 'Driver Performance')
```

Poles at this circuit:

```{r}
q=paste0('SELECT code AS Driver, COUNT(*) AS Poles FROM races r JOIN circuits c JOIN results rs JOIN drivers d WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND d.driverId=rs.driverId AND grid=1 AND circuitref="',cct$circuitRef,'" GROUP BY code DESC ORDER BY Poles DESC')
winners = dbGetQuery(ergastdb, q)
kable(winners, format='markdown',row.names = FALSE)
```

Individual winners at this circuit:

```{r}
q=paste0('SELECT code AS Driver, COUNT(*) AS Wins FROM races r JOIN circuits c JOIN results rs JOIN drivers d WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND d.driverId=rs.driverId AND position=1 AND circuitref="',cct$circuitRef,'" AND code IN (',driversThisYear_str,') GROUP BY code DESC ORDER BY Wins DESC')
winners = dbGetQuery(ergastdb, q)
kable(winners, format='markdown',row.names = FALSE)
```


Podiums at this circuit:

```{r}
q=paste0('SELECT code AS Driver, COUNT(*) AS Podiums FROM races r JOIN circuits c JOIN results rs JOIN drivers d WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND d.driverId=rs.driverId AND position IN (1,2,3) AND circuitref="',cct$circuitRef,'" AND code IN (',driversThisYear_str,') GROUP BY code DESC ORDER BY Podiums DESC')
winners = dbGetQuery(ergastdb, q)
kable(winners, format='markdown',row.names = FALSE)
```

```{r team_circuit_performance}

q=paste0('SELECT cn.name AS code, MAX(year) as until, SUM(points) s FROM races r JOIN circuits c JOIN results rs JOIN constructors cn WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND  cn.constructorId=rs.constructorId AND circuitref="',cct$circuitRef,'" GROUP BY cn.constructorId DESC ORDER BY s DESC LIMIT 5')
leaders = dbGetQuery(ergastdb, q)

q=paste0('SELECT year, cn.name AS code, position, points FROM races r JOIN circuits c JOIN results rs JOIN constructors cn WHERE r.circuitId=c.circuitId AND r.raceId=rs.raceId AND cn.constructorId=rs.constructorId AND circuitref="',cct$circuitRef,'" ORDER BY year')
res = dbGetQuery(ergastdb, q)


res = ddply(res, .(code), transform, totPts=cumsum(points)  )
champ_chart(res,leaders, 'Constructor Performance')
```

```{r}
# TO DO
#at circuit:
# pole sitters who win (conversion rate: pole-> win)
# pole sitters who made podium  (conversion rate: pole-> podium)
```

### Circuit Competitive Supertimes

*Supertimes* are often provided as a season wide performance metric for comparing performances over a season. At the manufacturer level, they are typically calculated for each manufacturer as the average of their fastest single lap recorded by the team at each race weekend expressed as a percentage of the fastest single lap overall.

We can also derive a reduced *competitive supertime* by basing the calculation on best laptime recorded across the qualifying and race sessions, omitting laptimes recorded in the practice sessions.

We can draw on the notion of supertimes to derive two simple measures for comparing team performances based on laptime:

- evolution of manufacturer competitive supertime for a circuit over the years;
- evolution of manufacturer competitive supertime for each circuit over the course of a season.

We can also produce driver performance metrics based on the competitive supertime of each driver.


```{r}
q=paste0('SELECT q1, q2, q3, fastestLapTime, cn.name, d.code, year FROM races r JOIN circuits c JOIN results rs JOIN constructors cn JOIN qualifying q JOIN drivers d WHERE r.circuitId=c.circuitId AND d.driverId=rs.driverId AND r.raceId=rs.raceId AND cn.constructorId=rs.constructorId AND q.raceId=r.raceId AND q.driverId=rs.driverId AND cn.name IN (',teamsThisYear_str,') AND circuitref="',cct$circuitRef,'" ORDER BY year')
st = dbGetQuery(ergastdb, q)


st = melt(st,
          id.vars = c("name","code", "year"),
          measure.vars = c("q1", "q2", "q3", "fastestLapTime"))

st['time'] = as.numeric(apply(st['value'], 1, timeInS))

st = ddply(st, .(year), transform, ntime = time/min(time, na.rm=TRUE))
stt = ddply(st,
            .(name,year),
            summarise, 
            stime = min(ntime, na.rm=TRUE),
            time =  min(time, na.rm=TRUE))

stt$label = as.factor(stt$name)
stt
```

```{r}
compSupertime = function(df, cct,typ='stime', labelSize=0.7, smooth=FALSE,
                         ylabel = 'Competitive Supertime (% of best)'){
  g=ggplot(data=df,aes_string(x='year',y=typ))
  if (smooth){g=g+stat_smooth(aes(colour=label),method = "lm", formula= y ~ x + I(x^2), se=FALSE, size=0.7)}
  else { g=g+geom_line(aes(colour=label))}
  g=g+ guides(colour=FALSE)+xlim(min(df$year),max(df$year)+1)
  #cex is label size
  g=g+geom_dl(aes(label = label, colour=label),method = list(cex = labelSize, dl.trans(x = x + .3), "last.bumpup"))
  g=g+labs(x=NULL,y=ylabel,
           title=paste0('F1 ', cct$race, ' - ','Competitive Supertimes',', ', min(stt$year), ' to ', max(df$year)),
           subtitle = paste0(cct$circuit,', ',cct$location,', ',cct$country),
           caption="Data from Ergast database, ergast.com/mrd")
  champChartTheme(g)
}
```

## Competitive Laptime Evolution

Evolution of best competitive laptime over the years, based on the best laptimes recorded in either qualifying or the race.

```{r bestCompetitiveTimeCircuit}
compSupertime(stt,cct,smooth=F,typ='time', ylabel='Best competitive laptime (s)')+ylim(88,95)
```


```{r supertimeCircuit}
#stt=stt[stt$year>2013,]
compSupertime(stt,cct,smooth=F)
```

```{r supertimeCircuitSmooth}
#stt=stt[stt$year>2013,]
compSupertime(stt,cct,smooth=T)
```



```{r driverSupertimeCircuit}

dst = ddply(st, .(code,year), summarise, stime = min(ntime, na.rm=TRUE))

dst$label=as.factor(dst$code)
#dst=dst[dst$year>2013,]

dstt =  ddply(dst, .(code), summarise,
             avstime=mean(stime, na.rm=TRUE),
             races=sum(year>0))

best=head(dstt[order(desc(dstt$races),dstt$avstime),],5)$code

compSupertime(dst,cct,labelSize=0.4) 
```

```{r driverSupertimeCircuitSmooth}
compSupertime(dst,cct,labelSize=0.4, smooth=T) 
```

```{r}
dstt=dstt[dstt$code %in% driversThisYear,]
#kable(dstt[order(desc(dstt$races),dstt$avstime),c('code','avstime')], format='markdown',row.names = FALSE)
kable(dstt[order(desc(dstt$races),dstt$avstime),], format='markdown',row.names = FALSE)

```

```{r}
dstt=dstt[dstt$code %in% driversThisYear,]
#kable(dstt[order(desc(dstt$races),dstt$avstime),c('code','avstime')], format='markdown',row.names = FALSE)
kable(dstt[order(dstt$avstime),], format='markdown',row.names = FALSE)

```



